### 2018-07-26 Main.Rmd

---
title: "Replication of Hills 2009"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Background

...

# Networks

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
library(ForceAtlas2)

source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)
```

## Nouns

We begin with roughly the same data used by Hills. *(TODO fix, get same data)*

```{r warning=FALSE}
# We begin by importing data from wordbank. Age of aquisition (in months, it starts with 16, 17,..) is that at which 50% of children can produce the word. We also do naive triming at this point, this means we are ignoring homophone/polysemy (chiken (food), chiken (animals), etc,.. but we will have to be careful in our choices if we are preparing later for publication.

full_vocab <- make_vocab_dataframe(lang="English (American)",
                              lang_form = "WS",
                              lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age)
```

The initial vocabulary contains `r nrow(full_vocab)` (compared to Hills' 312).

## Perceptual & Conceptual Features

Features taken from McRae et al. (2005), in which features were collected for 541 nouns from 725 adults, with 30 adults providing  14 features for each noun.

```{r warning=FALSE}
features <- load_features()

hills_vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter(uni_lemma %in% features$Concept) # filter to words with features
```

To begin our analysis, we only consider the words for which feature data are available, an overlap of `r nrow(vocab)` in our data compared to Hills' 130.

## Word2Vec

## Build Networks

Using vocabulary obtained by 30 months and the same McRae features as Hills, we construct networks from perceptual and conceptual features. We track the number of features that generate each edge.

```{r message=FALSE}
# perceptual and conceptual McRae features
perceptual_features = c("sound", "taste", "smell", "tactile",
                        "visual-form_and_surface", "visual-motion", "visual-color")
conceptual_features = c("function")

# count shared features between words in vocabulary
perceptual_links <- make_feature_links(vocab = hills_vocab,
                                       feature_types = perceptual_features) %>%
  rename(perceptual = shared)

conceptual_links <- make_feature_links(vocab = hills_vocab,
                                       feature_types = conceptual_features) %>%
  rename(conceptual = shared)

# get word2vec distance between words in vocabulary
w2v_links <- make_w2v_links(vocab = hills_vocab) #TODO still problem w/ vocab matching w2v data

links <- perceptual_links %>%
  full_join(conceptual_links, by=c("from", "to")) %>%
  full_join(w2v_links, by=c("from", "to"))

full_network <- graph_from_data_frame(links, vertices = hills_vocab, directed = FALSE) %>%
  as_tbl_graph()
```

# Results

## The Full Network

### Network statistics and small-world analyses

```{r}
# testing dense layout
source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)

plot_network(full_network %E>% filter(w2v_similarity >= 0.5), remove_isolated = TRUE, labels=TRUE, frame=TRUE)
```

The full network is shown in the following figure, for w = 1, 2, 3, and 4.

```{r out.width = '100%'}
full_networks = list()
for (w in 1:4) {
  full_networks[[w]] = full_network %E>%
    filter(shared >= w)
}

#par(mfcol=c(2, 2), mar=c(0.5,0.5,0.5,0.5))
for (network in full_networks) {
  plot_network(network, remove_isolated = FALSE, labels=TRUE, frame=TRUE)
}
```

## Basic Vocab and Feature Properties

<!-- ## Basic Network Properties -->

<!-- Some basic properties of networks built from perceptual and conceptual features. -->
<!-- ```{r, results='asis'} -->
<!-- max_w <- 4 -->

<!-- network_properties <- function(networks) { -->
<!--   return( tibble( -->
<!--   name = map_chr(networks, ~ .x$name), -->
<!--   type = map_chr(networks, ~ .x$type), -->
<!--   age = map_dbl(networks, ~ .x$age) %>% -->
<!--     as.integer(), -->
<!--   w = map_dbl(networks, ~ .x$w) %>% -->
<!--     as.integer(), -->
<!--   vertices = map_int(networks, vcount), -->
<!--   edges = map_dbl(networks, ecount) %>% -->
<!--     as.integer(), # for some reason igraph ecount returns double -->
<!--   avg_degree = map(networks, degree) %>% -->
<!--     map_dbl(mean), -->
<!--   avg_shortest_path = map_dbl(networks, mean_distance), -->
<!--   diameter = map_dbl(networks, diameter) %>% -->
<!--     as.integer(), -->
<!--   clustering_coefficient = map_dbl(networks, transitivity), -->
<!--   density = edges / (vertices * (vertices - 1) / 2), -->
<!--   avg_shortest_path_random = 0, # either estimate or calculate values for random nets -->
<!--   diameter_random = 0, -->
<!--   clustering_coefficient_random = density, -->
<!--   isolates = map_dbl(networks, function (g) sum(degree(g) == 0)), -->
<!--   components = map_dbl(networks, count_components) %>% -->
<!--     as.integer() - isolates)) -->
<!-- } -->

<!-- full_networks <- list() -->
<!-- perceptual_networks <- list() -->
<!-- conceptual_networks <- list() -->
<!-- for (w in 1:max_w) { -->
<!--   full_networks[[w]] <- full_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->

<!--   perceptual_networks[[w]] <- perceptual_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->

<!--   conceptual_networks[[w]] <- conceptual_network %>% -->
<!--     activate(what = edges) %>% -->
<!--     filter(shared >= w) %>% -->
<!--     set.graph.attribute("w", w) -->
<!-- } -->

<!-- networks <- full_networks %>% -->
<!--   append(perceptual_networks) %>% -->
<!--   append(conceptual_networks) -->

<!-- our_network_properties <- network_properties(networks) -->

<!-- our_network_properties %>% -->
<!--   select(type, w, vertices, edges, avg_degree, diameter, avg_shortest_path, clustering_coefficient) %>% -->
<!--   kable(format="html", -->
<!--         escape = FALSE, -->
<!--         col.names = c("Type", "w", "V", "E", "Average degree", "Diameter", "Average length", "Clustering coefficient"), -->
<!--         digits = 2, -->
<!--         align = rep("l", 8)) %>% -->
<!--   kable_styling(full_width = F) %>% -->
<!--   collapse_rows(columns = 1:3, valign = "top") -->
<!-- ``` -->

<!-- We can compare these networks to those from Hills, noting that we have a different set of words/nodes but that the induced subgraph of those words shared between our analysis and that of Hills should be the same. -->
<!-- ```{r} -->
<!-- hills_network_properties <- tibble( # table from Hills 2009 -->
<!--   type = c(rep("all", times=4), rep("perceptual", times=4), rep("conceptual", 4)), -->
<!--   w = c(1:max_w, 1:max_w, 1:max_w), -->
<!--   clustering_coefficient = c(0.55, 0.54, 0.58, 0.60, 0.54, 0.54, 0.62, 0.62, 0.88, 0.59, 0.38, 1.00), -->
<!--   avg_shortest_path = c(1.72, 2.57, 4.96, 3.22, 1.76, 2.69, 5.02, 2.55, 3.08, 1.59, 5.02, 1.00), -->
<!--   density = c(0.29, 0.10, 0.04, 0.02, 0.27, 0.08, 0.03, 0.01, 0.05, 0.01, 0.0, 0.0), -->
<!--   components = c(1, 1, 4, 10, 1, 2, 8, 11, 7, 14, 7, 2), -->
<!--   isolates = c(0, 6, 20, 47, 0, 10, 31, 69, 33, 81, 113, 126) -->
<!-- ) -->

<!-- # convert contents to text -->
<!-- display_our_properties <- our_network_properties %>% mutate_all(funs(format(., digits=2, trim=TRUE))) -->
<!-- display_hills_properties <- hills_network_properties %>% mutate_all(funs(paste("(", ., ")", sep=''))) -->

<!-- # combine text tables for display -->
<!-- display_properties <- bind_rows(display_our_properties %>% mutate(row = 1:n()), -->
<!--                                 display_hills_properties %>% mutate(row = 1:n())) %>% -->
<!--   group_by(row) %>% -->
<!--   summarise_if(is.character, funs(paste(., collapse=" "))) %>% -->
<!--   arrange(row) -->

<!-- # display combined properties -->
<!-- select(hills_network_properties, type, w) %>% -->
<!--   bind_cols(select(display_properties, clustering_coefficient, avg_shortest_path, -->
<!--                    density, components, isolates)) %>% -->
<!--   kable(format="html", -->
<!--         escape = FALSE, -->
<!--         col.names = c("Type", "w", "Clustering Coefficient", "Average length", "Density", "Components", "Isolates"), -->
<!--         align = rep("l", 7)) %>% -->
<!--   kable_styling(full_width = F) %>% -->
<!--   collapse_rows(columns = 1:2, valign = "top") -->
<!-- ``` -->

<!-- # Clusters -->

<!-- ## Visualize Networks -->
<!-- ```{r out.width = '100%'} -->
<!-- networks <- perceptual_networks[2] -->
<!-- ``` -->

<!-- ## Clique Percolation -->

<!-- Initial Clique Analysis -->
<!-- ```{r out.width = '100%'} -->
<!-- trimmed_networks <- map(networks, function (g) delete.vertices(g, degree(g)==0)) -->

<!-- # cl <- makeCluster(3) -->
<!-- # registerDoParallel(cl) -->
<!-- # getDoParWorkers() -->
<!-- ## Plot with the new colors -->
<!-- for (net in networks) { -->
<!--   for (k in 3:4) { -->
<!--     clusters <- clique.community.opt(net,k) -->
<!--     plot_network(net, clusters = clusters, labels=FALSE) -->
<!--   } -->
<!-- } -->
<!-- # stopCluster(cl) -->
<!-- ``` -->

<!-- Dynamic Visualization -->
<!-- ```{r} -->
<!-- conceptual_clusters <- membership(cluster_walktrap(conceptual_network)) -->

<!-- # Convert to object suitable for networkD3 -->
<!-- conceptual_d3 <- igraph_to_networkD3(conceptual_network, group = conceptual_clusters) -->

<!-- # Create force directed network plot -->
<!-- forceNetwork(Links = conceptual_d3$links, Nodes = conceptual_d3$nodes, -->
<!--              Source = 'source', Target = 'target', NodeID = 'name', -->
<!--              Group = 'group') -->
<!-- ``` -->