### 2018-08-28 HillsComparison.Rmd

---
title: "Comparison Word2Vec Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Preparation

This analysis requires a working copy of CFinder and the CFinder license in the same directory (necessary for running CFinder). The following must point to the command line executable:

```{r}
cfinder_program <- "./CFinder_commandline_mac"
```

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
library(viridis)

source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)
```

## Nouns

We begin with roughly the same data used by Hills. *(TODO fix, get same data)*

```{r warning=FALSE}
# We begin by importing data from wordbank. Age of aquisition (in months, it starts with 16, 17,..) is that at which 50% of children can produce the word. We also do naive triming at this point, this means we are ignoring homophone/polysemy (chiken (food), chiken (animals), etc,.. but we will have to be careful in our choices if we are preparing later for publication.

full_vocab <- make_vocab_dataframe(lang="English (American)",
                                   lang_form = "WS",
                                   lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age, category)
```

The initial vocabulary contains `r nrow(full_vocab)` (compared to Hills' 312).

## Features and Word2Vec

Features taken from McRae et al. (2005), in which features were collected for 541 nouns from 725 adults, with 30 adults providing 14 features for each noun. Word2Vec similarities are from child directed speach (source Abdellah).

```{r warning=FALSE, message=FALSE}
vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter_to_features() %>%
  filter_to_w2v()
```

Filtering to this data gives a vocab of `r nrow(vocab)`.

## Build Network

Using this vocabulary, we construct a network with edge attributes for features, and for word2vec similarity.

```{r message=FALSE}
# perceptual and conceptual McRae feature types
perceptual_features = c("sound", "taste", "smell", "tactile",
                        "visual-form_and_surface", "visual-motion", "visual-color")
conceptual_features = c("function")

# count shared features between words in vocabulary
feature_links <- make_feature_links(vocab = vocab,
                                    feature_types = union(perceptual_features,
                                                          conceptual_features))

# get word2vec distance between words in vocabulary
w2v_links <- make_w2v_links(vocab = vocab) #TODO still problem w/ vocab matching w2v data

links <- full_join(feature_links, w2v_links, by=c("from", "to"))

network <- graph_from_data_frame(links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph() %E>%
  mutate(shared = replace_na(shared, as.integer(0))) # is this replacement necessary?
```

# Results

## Word2Vec vs Feature Network

```{r out.width = '25%', fig.asp=1, fig.show = 'hold'}
par(oma=c(0,0,0.2,0))

ecounts <- list()
for (w in 1:4) {
  net <- network %E>% filter(shared >= w)
  ecounts[[length(ecounts) + 1]] <- ecount(net)
  layout <- layout_with_fr(net)
  plot_vocab_network(net,
                     plot_title = title(str_glue("Features w={w}"), cex.main = 2),
                     frame=FALSE, layout = layout, labels="")
}

# find thresholds that give matching edge counts
thresholds <- list()
for (i in seq_along(ecounts)) {
  e <- ecounts[[i]]
  curr_pow <- 1
  curr_e <- 0
  curr_t <- 1
  while (as.integer(curr_e) != as.integer(e)) {
    if (curr_e < e) {
      curr_t <- curr_t - (1/2) ^ curr_pow
    } else {
      curr_t <- curr_t + (1/2) ^ curr_pow
    }
    curr_e <- ecount(network %E>% filter(w2v_similarity >= curr_t))
    curr_pow <- curr_pow + 1
  }
  thresholds[[length(thresholds) + 1]] <- curr_t
}

for (w in thresholds) {
  net <- network %E>% filter(w2v_similarity >= w)
  layout <- layout_with_fr(net)
  plot_vocab_network(net, plot_title = title(str_glue("Word2Vec w={format(w,digits=3)}"), cex.main = 2),
                     frame=FALSE, layout = layout, labels="")
}

```

```{r}
network_properties <- function(network) {
  return( tibble(
    type = network$type,
    w = network$w,
    avg_distance = mean_distance(network),
    clustering_coef = transitivity(network),
    isolates = sum(degree(network) == 0),
    components = count_components(network) - isolates
  ))
}

feature_networks <- map(1:4, function(w) {
  network %>%
    filter(shared >= w) %>%
    set.graph.attribute("type", "Feature") %>%
    set.graph.attribute("w",w)})
w2v_networks <- map(thresholds, function(w) {
  network %>%
    filter(w2v_similarity >= w) %>%
    set.graph.attribute("type", "Word2Vec") %>%
    set.graph.attribute("w",w)}
)

feature_properties <- bind_rows(map(feature_networks, network_properties))
w2v_properties <- bind_rows(map(w2v_networks, network_properties))

bind_rows(feature_properties, w2v_properties) %>%
  kable(format="html",
        escape = FALSE,
        col.names = c("Type", "w",
                      "Clustering Coefficient",
                      "Average length",
                      "Components",
                      "Isolates"),
        align = rep("l",6),
        digits = 2) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, valign = "top")
```

## Clique Percolation

We can now compare clique percolation on feature and word2vec networks.

```{r}
best_count <-  0
for (w in 1:5) {
  result <- cluster_cfinder(network %E>% filter(shared >= w),
                            cfinder_program = cfinder_program)
  clusters <- result$clusters
  k <- result$k
  if (length(clusters) > best_count) {
    best_w <- w
    best_k <- k
    best_count <- length(clusters)
    best_clusters <- clusters
  }
}

title <- str_glue("Feature Network Clique Percolation Clusters, w = {best_w} and k = {best_k}")
kable_clusters(best_clusters, title)
```

We can compare this to the corresponding clustering on the word2vec network with the same density, using the same parameters.

```{r}
clusters <- cluster_cfinder(network %E>%
                              filter(w2v_similarity >= thresholds[[best_w]]),
                            k = best_k,
                            cfinder_program = cfinder_program)$clusters

title <- str_glue("Word2Vec Network Clique Percolation Clusters, w = {format(thresholds[[best_w]], digits=2)} and k = {best_k}")
kable_clusters(clusters, title)
```

We can also search for parameters for the word2vec network using Hills' approach. Note that
w is given a lower bound of 0.25 as below this the network is too dense for cfinder to compute the clusters in a minute or less.

```{r}
best_count <-  0
for (w in seq(0.25,0.9,0.05)) {
  result <- cluster_cfinder(network %E>% filter(w2v_similarity >= w),
                            cfinder_program = cfinder_program)
  clusters <- result$clusters
  k <- result$k
  if (length(clusters) > best_count) {
    best_w <- w
    best_k <- k
    best_count <- length(clusters)
    best_clusters <- clusters
  }
}

title <- str_glue("Word2Vec Network Clique Percolation Clusters, w = {best_w} and k = {best_k}")
kable_clusters(best_clusters, title)
```

## Other Clustering Algorithms

We can cluster both the feature and word2vec networks using other clustering algorithms.

### Walktrap

```{r}
clusters <- cluster_walktrap(network, weights = E(network)$shared) # NA weights are treated as 0
kable_clusters(clusters, title = "Feature Network Walktrap Clusters")
```

```{r}
clusters <- cluster_walktrap(network, weights = E(network)$w2v_similarity)
kable_clusters(clusters, title = "Word2Vec Network Walktrap Clusters")
```

### Louvain

```{r}
clusters <- cluster_louvain(network, weights = E(network)$shared) # NA weights are treated as 0
kable_clusters(clusters, title = "Feature Network Louvain Clusters")
```

```{r}
clusters <- cluster_louvain(network, weights = E(network)$w2v_similarity)
kable_clusters(clusters, title = "Word2Vec Network Louvain Clusters")
```

### Spinglass

```{r}
clusters <- cluster_spinglass(network, weights = E(network)$shared) # NA weights are treated as 0
kable_clusters(clusters, title = "Feature Network Spinglass Clusters")
```

```{r}
clusters <- cluster_spinglass(network, weights = E(network)$w2v_similarity)
kable_clusters(clusters, title = "Word2Vec Network Spinglass Clusters")
```