### 2018-08-28 Word2Vec.Rmd

---
title: "Word2Vec Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Networks

This report explores the clustering of nouns aquired by 16 - 30 month old children using word2vec similarity, and the dynamics of those clusters.

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
library(viridis)
library(riverplot) # install from github at ischeinfeld/riverplot to hide debug messages in CRAN package
library(ggm)
library(networkD3)

source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)
```

## Nouns

We begin by importing data from wordbank. Age of aquisition (in months, it starts with 16, 17,..) is that at which 50% of children can produce the word. We also do naive triming at this point, this means we are ignoring homophone/polysemy (chiken (food), chiken (animals), etc,.. but we will have to be careful in our choices if we are preparing later for publication.

```{r warning=FALSE}
full_vocab <- make_vocab_dataframe(lang="English (American)",
                                   lang_form = "WS",
                                   lex_class = "nouns") %>%
  trim_all_unilemma() %>%
  trim_all_definition() %>%
  arrange(age) %>%            # keep first age for duplicates
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(uni_lemma, definition, age, category)
```

The initial vocabulary contains `r nrow(full_vocab)` words.

## Word2Vec

Features taken from McRae et al. (2005), in which features were collected for 541 nouns from 725 adults, with 30 adults providing 14 features for each noun. Word2Vec similarities are from child directed speach (source Abdellah).

```{r warning=FALSE, message=FALSE}
vocab <- full_vocab %>%
  filter(age <= 30) %>%
  filter_to_w2v()
```

To begin our analysis, we only consider the words for which word2vec data are available, an overlap of `r nrow(vocab)` in our data. The missing words are `r setdiff(full_vocab$uni_lemma, vocab$uni_lemma)`.

## Build Network

We construct a network with an edge attribute for word2vec similarity.

```{r message=FALSE}
w2v_links <- make_w2v_links(vocab = vocab) #TODO still problem w/ vocab matching w2v data

network <- graph_from_data_frame(w2v_links, vertices = vocab, directed = FALSE) %>%
  as_tbl_graph() %E>%
  mutate(weight = w2v_similarity) # layout and clustering algorithms automatically use the weight attribute
```

# Analysis

## Layout the Network

First we generate a display layout for the network, based on the word2vec data.

```{r out.width = '100%', fig.asp = 1, dpi=300}
if (FALSE) {
  layout <- layout_network(network)
  saveRDS(layout, file = "layout.rds")
} else {
  layout <- readRDS(file = "layout.rds")
}

network <- network %N>%
  mutate(x = layout[,1]) %>%
  mutate(y = layout[,2])

plot_vocab_network(network,
                   frame=FALSE,
                   plot_title=title("Word2Vec Network",cex.main=1),
                   labels="",
                   edge_color = rgb(0,0,0,alpha = (E(network)$weight)^(5/2)))

```

## Clustering algorithms {.tabset}

### Walktrap

```{r out.width='33%', fig.asp=1, dpi=300, results='asis'}
par(oma=c(0,0,0.2,0))

for (age_limit in seq(16, 30, by=2)) {

  net <- network %N>% filter(age <= age_limit)
  clusters <- cluster_walktrap(net,
                               weights = E(net)$w2v_similarity)

  filename <- str_glue("images/walktrap_network_{age_limit}.png")
  png(filename=filename)
  plot_vocab_network(network %N>%
                       filter(age <= age_limit),
                     clusters = clusters$membership,
                     frame = FALSE,
                     labels = "",
                     edge_color = rgb(0,0,0,alpha = (E(network)$weight)^(5/2)))
  dev.off()

  cat("<table style='width:100%'><col width='30%'><col width='70%'><tr>")

  cat("<td valign='top'>")

  cat(str_glue('<img src="{filename}">'))

  cat("</td><td valign='top'>")

  print(kable_clusters(clusters,
                 title=str_glue("Age = {age_limit} Months")))
  cat("</tr></table>")

}
```

#### River Plot

```{r}
subgraphs_by_age <- get_subgraphs_by_age(network, cluster_walktrap)
```

```{r out.width='100%'}
#out.width='100%', fig.asp=1, dpi=300
river <- cluster_river(subgraphs_by_age = subgraphs_by_age)

plot(river, gravity = "center", plot_area = 0.8)
```

```{r}
suppressWarnings(precursors_graph(subgraphs_by_age, 0)) %>% plot_river_network()
```