### 2018-11-30 languages.Rmd
similarity based growth

---
title: "Comparing order of acquisition cross-linguistically"
author: "Abdellah Fourtassi"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---
# Preparation

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(langcog) # useful things used by Language and Cognition Lab
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(broom) # converts R function outputs to tidy data frames
library(igraph) # network analysis and visualization
library(knitr) # general-purpose tool for dynamic report generation
library(kableExtra) # table formatting
library(lemon) # functions for working w/ ggplot2 and knitr extensions
library(tibble) # modern dataframes
library(tidygraph)
library(doParallel)
library(networkD3)
#library(ForceAtlas2)
library(boot)
library(ggthemes)

source(paste(getwd(),"/helpers/all_helper.r",sep = ""), chdir = T)
```

Get cross-linguistic data from Wordbank
```{r}

languages = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish")

admins <- get_administration_data() %>%
  select(data_id, age, language, form) %>%
  filter(language %in% languages)

items <- get_item_data() %>%
  filter(type == "word", lexical_class == "nouns") %>%
  filter(language %in% languages)

items_by_inst <- split(items, paste(items$language, items$form, sep = "_"))

get_inst_data <- function(inst_items) {
  inst_lang <- unique(inst_items$language)
  inst_form <- unique(inst_items$form)
  inst_admins <- filter(admins, language == inst_lang, form == inst_form)
  get_instrument_data(language = inst_lang ,
                      form = inst_form,
                      administrations = inst_admins,
                      items = inst_items$item_id,
                      iteminfo = inst_items
                      ) %>%
    filter(!is.na(age)) %>%
    mutate(produces = !is.na(value) & value == "produces",
           understands = !is.na(value) & (value == "understands" | value == "produces")) %>%
    select(-value) %>%
    gather(measure, value, produces, understands) %>%
    filter((measure == "understands" & form == "WG") | (measure == "produces" & form == "WS") ) %>%
    mutate(language = inst_lang,
           form = inst_form)

}

data_raw <- map(items_by_inst, get_inst_data)

data_all <- bind_rows(data_raw) %>%
  rename(item = num_item_id) %>%
  group_by(language, form, measure,
             lexical_category, lexical_class, uni_lemma,  item, definition, 
             age) %>%
  summarise(num_true = sum(value, na.rm = TRUE),
              num_false = n() - num_true,
              prop = mean(value, na.rm = TRUE))

#feather::write_feather(data_all, "saved_data/data_all.feather")

#I am taking out measure to keep only produce
data_all <- feather::read_feather("saved_data/data_all.feather")

data_all$language <- plyr::mapvalues(data_all$language, 
                                 from = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", "Russian", "Spanish", "Swedish", "Turkish"))

```

Calculate the age of acquisition
```{r}

fit_inst_measure_uni <- function(inst_measure_uni_data) {
  ages <- min(inst_measure_uni_data$age):max(inst_measure_uni_data$age)
  model <- glm(cbind(num_true, num_false) ~ age, family = "binomial",
               data = inst_measure_uni_data)
  fit <- predict(model, newdata = data.frame(age = ages), se.fit = TRUE)
  aoa <- -model$coefficients[["(Intercept)"]] / model$coefficients[["age"]]
  constants <- inst_measure_uni_data %>%
    ungroup()%>%
    select(language, form, measure, lexical_category, lexical_class, uni_lemma, item, definition) %>%
    distinct()

  props <- inst_measure_uni_data %>%
    ungroup() %>%
    select(age, prop)

  data.frame(age = ages,
             fit_prop = inv.logit(fit$fit),
             fit_se = fit$se.fit,
             aoa = aoa, language = constants$language,
             form = constants$form,
             measure = constants$measure,
             uni_lemma = constants$uni_lemma,
             item = constants$item,
             definition = constants$definition) %>%
    left_join(props)
}

list_by_item <- data_all %>%
  # make this filtering at the beginning, not here
  split(paste(.$language, .$form, .$measure, .$item))

data_aoa <- map(list_by_item, fit_inst_measure_uni) %>%
  bind_rows()

data_aoa <- data_aoa %>%
  select(language, form, measure, uni_lemma, item, definition, aoa) %>%
  distinct()

#feather::write_feather(data_aoa, "saved_data/data_aoa.feather")

data_aoa <- feather::read_feather("saved_data/data_aoa.feather")

#I am taking out measure to keep only produce
#data_aoa2 <- feather::read_feather("saved_data/aoa_data.feather") %>%
#  rename(definition = defintion) %>%
#  select(language, form, measure, uni_lemma, item, definition, aoa) %>%
#  distinct()

```

Colculate the correlation
```{r}

#Correlation between comprehension and production in each language
#For each pairwise comparision, compute correlation using values unilemmas which are avialable in both languages

#I should have a double loop?
#For each pair of languages, select the unilemmas and their aoa, rank the unilemmas based on aoa, and calcualte the correlation

# rank unilemma asa function of AoA

data_prod <- data_aoa %>%
  filter(measure =='produces')

data_by_lang <- split(data_prod, data_prod$language)

pair_data <- function(lang1){

  pair_cor_data <- function (lang2){

  lang1_unilemmas <- unique(lang1$uni_lemma)
  lang2_unilemmas <- unique(lang2$uni_lemma)
  uni_intersect <-  intersect(lang1_unilemmas, lang2_unilemmas)

  data_lang1 <- lang1 %>%
    filter(!is.na(uni_lemma),
           uni_lemma %in% uni_intersect) %>%
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa1=mean(aoa))

  data_lang2 <- lang2 %>%
    filter(!is.na(uni_lemma),
           uni_lemma %in% uni_intersect) %>%
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa2=mean(aoa))

  both <- data_lang1 %>%
    left_join(data_lang2)

  mycorr <- cor.test(rank(both$aoa1), rank(both$aoa2), method = 'kendall')
  val=mycorr$estimate
  p=mycorr$p.value
  language1 = unique(lang1$language) 
  language2 = unique(lang2$language)

  data.frame(language1=language1, language2=language2, tau=val, sig = p)

}
  map(data_by_lang, pair_cor_data) %>%
    bind_rows()

}

pair_random <- function(lang1){

  pair_cor_rand <- function (lang2){

  data_lang1 <- lang1 %>%
    filter(!is.na(uni_lemma)) %>%
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa1=mean(aoa))

  order_lang1 <- rank(data_lang1$aoa1)
  order_random <- sample(order_lang1, size = length(order_lang1), replace = FALSE)

  mycorr <- cor.test(order_lang1, order_random, method = 'kendall')

  val=mycorr$estimate
  p=mycorr$p.value
  language1 = unique(lang1$language) 
  language2 = unique(lang2$language)

  data.frame(language1=language1, language2=language2, tau=val, sig = p)

}
  map(data_by_lang, pair_cor_rand) %>%
    bind_rows()

}

cor_data <- map(data_by_lang, pair_data) %>%
  bind_rows() %>%
  filter(language1!=language2) %>%
  mutate(data="languages")

cor_random <- map(data_by_lang, pair_random) %>%
  bind_rows() %>%
  mutate(data="random")

```

similarity-based learning mechanism
```{r}

unilemmas <- (data_by_lang[[3]] %>% #English (I should recode this to be more explcit)
    filter(!is.na(uni_lemma)) %>%
    trim_all_unilemma() %>%
    select(uni_lemma) %>%
    #(uni_lemma))$uni_lemma
    unique())$uni_lemma

#get the data frame for w2v
w2v <- load_w2v() %>%
  filter(from %in% unilemmas & to  %in% unilemmas)

#Transfom the data fram into adjencency matrix
g <- graph.data.frame(w2v, directed=FALSE)
  #matrix_sim <- get.adjacency(g, attr="w2v_similarity", sparse=FALSE)
  matrix_sim <- get.adjacency(g, attr="w2v_similarity")

vocab <- names(matrix_sim[word_0, ])
#Now generate a sequence based on the first word

sim_learning <- function() {
  seq_gen <- function(word_t0, sequence_t0) {

    if (length(sequence_t0) == length(matrix_sim[word_t0, ])) {
      return (sequence_t0)
    } else {
      #generate similarity values of all words to word_t
      all_words <- matrix_sim[word_t0, ]
      #Obtain indices of the words already known at time t (present in in sequence_t)
      indices <- which(names(all_words) %in% sequence_t0)
      #generate candidates for word_t+1 by subseting known words from all words
      candidates <- all_words[-indices]
      #the next word (i.e., word_t+1) is the most simialr to the last word acquired ()
      index_t1 <- which.max(candidates)
      word_t1 <- names(index_t1)
      sequence_t1 <- c(sequence_t0, word_t1)

      return(seq_gen(word_t1, sequence_t1))
    }
  }

  word_0 <- sample(vocab, 1) # this is the only random compoenent
  sequence_0 <- c(word_0)
  #seq_gen(word_0, sequence_0)
  return(seq_gen(word_0, sequence_0))
}

pair_sim <- function(lang1){
  pair_cor_sim <- function (lang2){

  #Orering based on word similarity

  #Real ordering in the language 
  data_lang1 <- lang1 %>%
    filter(!is.na(uni_lemma)) %>%
    trim_all_unilemma() %>%
    filter(uni_lemma %in% vocab) %>% #Intersection with w2v data
    select(uni_lemma, aoa) %>%
    group_by(uni_lemma) %>%
    summarise(aoa1=mean(aoa))

  #Orering based on word similarity
  sim_list <- sim_learning()
  sim_data <- data.frame(uni_lemma=sim_list, orderSim = seq(1, length(sim_list)))

  #Combine both
  both_orders <- data_lang1 %>%
    left_join(sim_data) %>%
    filter(!is.na(orderSim),
           !is.na(aoa1))

  order_lang1 <- rank(both_orders$aoa1)
  order_sim <- rank(both_orders$orderSim)

  mycorr <- cor.test(order_lang1, order_sim, method = 'kendall')

  val=mycorr$estimate
  p=mycorr$p.value
  language1 = unique(lang1$language) 
  language2 = unique(lang2$language)

  data.frame(language1=language1, language2=language2, tau=val, sig = p)

}
  map(data_by_lang, pair_cor_sim) %>%
    bind_rows()

}

cor_sim <- map(data_by_lang, pair_sim) %>%
  bind_rows() %>%
  mutate(data="sim")

cor_all <- cor_data %>%
  bind_rows(cor_random) %>%
  bind_rows(cor_sim)

```

plots
```{r}

ggplot(cor_all, 
      aes(x = tau, fill=data)) +
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
                 alpha=0.5,
                 binwidth=0.05)+
  #scale_x_log10() +
  theme_few() + 
  theme(aspect.ratio = 0.7, legend.title = element_text(size=8)) +  #facet_grid(Segmentation ~ language)+
  scale_y_continuous(labels = scales::percent)+
  xlab("pairwise correlation") +ylab("Count")

```