### 2019-01-28 conceptNet-cogsci/Cogsci.Rmd
copied cogsci to folder

---
title: "CogSci Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---

# Setup

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(igraph) # network analysis and visualization
library(tibble) # modern dataframes
library(tidygraph)
library(matrixStats)
library(memoise)

source("helpers/wb_helper.r")
source("helpers/word2vec_helper.r")
source("helpers/network_helper.r")
source("helpers/random_helper.r")
source("helpers/dynamics_helper.r")
source("helpers/statistics_helper.r")
```

## Nouns

We begin by importing data from wordbank. Age of aquisition (in months, it starts with 16, 17,..) is that at which 50% of children can produce the word. We also do naive triming at this point, this means we are ignoring homophone/polysemy (chiken (food), chiken (animals), etc,.. but we will have to be careful in our choices if we are preparing later for publication.

```{r warning=FALSE}
get_language_vocab_ <- function(lang) {
  return(make_vocab_dataframe(lang=lang,
                              lang_form = "WS",
                              lex_class = "nouns") %>%
           trim_all_unilemma() %>% # TODO check trimming
           trim_all_definition() %>%
           arrange(age) %>%            # keep first age for duplicates
           distinct(uni_lemma, .keep_all = TRUE) %>%
           select(uni_lemma, definition, age, category))
}

get_language_vocab <- memoize(get_language_vocab_) # memoize to prevent re-downloading dataset TODO check if this works
```

##

## Build Word2Vec Network Evolution

```{r message=FALSE}
clusters_evolution <- function(vocab, n_clusters, min_age, max_age, randomization) {

  if(randomization == "none") {
    # do nothing
  } else if(randomization == "random_aoa") {
    vocab <- randomize_aoa(vocab) # TODO handle randomization, write "randomize_aoa"" in random_helper.r
  } else if(randomization == "to_nearest_aoa") {
    vocab <- randomize_aoa_to_nearest(vocab)
  } else if(randomization == "within_last_clustering") {
    vocab <- randomize_within_last_clustering(vocab, n_clusters)
  } else {
    stop("Randomization argument incorrect.")
  }

  vocab_network <- vocab_to_w2v_network(vocab) # drops words w/o w2v data, NOTE memoizes w2v data

  # list of networks at each age
  net_evo <- min_age:max_age %>%
    map(~ filter(activate(vocab_network, nodes), age <= .))

  # list of vocabulaires at each age
  vocab_evo <- net_evo %>%
    map(V)

  # list of cluster membership index vectors at each age
  membership_evo <- net_evo %>%
    map(w2v_waltrap) %>%
    map(~ cut_at(., no = n_clusters))

  # list of word clusters at each age
  clusters_evo <- map2(vocab_evo, membership_evo, build_clusters)

  return(clusters_evo)
}
```

# Analysis

## Measure Clustering Dynamics

```{r}
# language, n_clusters, age range, random samples, dynamics measure -> analysis
# example measure is function(clusters_evo) {pair_measures_to_last(clusters_evo, measure)}
run_measure <- function(clusters_evo, measure) {
  if (measure == "pair_precision_to_last") {
    return(pair_measures_to_last(clusters_evo, pair_precision))
  } else if (measure == "pair_recall_to_last") {
    return(pair_measures_to_last(clusters_evo, pair_recall))
  } else if (measure == "pair_f1_to_last") {
    return(pair_measures_to_last(clusters_evo, pair_f1))
  } else if (measure == "min_max_size_ratio") {
    return(map_dbl(clusters_evo, min_max_size_ratio))
  } else if (measure == "modularity") {
    return(map_dbl(clusters_evo, min_max_size_ratio))
  } else {
    stop("Incorrect Dynamics Measure Name")
  }
}

# -----------------------------------------------------------------------------------------------
# given a language and # of clusters, the dynamics measures and randomizations are calculated
# together to avoid repeatedly sampling clusterings
# -----------------------------------------------------------------------------------------------

measure_cluster_dynamics <- function(language, n_clusters, measures, randomizations, samples) {
    vocab <- get_language_vocab(language)

    # measurement schema : age, measure, randomization, sample, value
    measurements <- list() # list of measurement tibbles

    # true vocabulary sizes (full vocabulary size at each month)
    measurements[["true_vocab_sizes"]] <-
      tibble(age = min(vocab$age):max(vocab$age),
             measure = "true_vocab_size",
             randomization = NA, # same for all randomizations
             sample = NA,
             value = min(vocab$age):max(vocab$age) %>%
               map(~ filter(vocab, age <= .)) %>%
               map(nrow) %>%
               unlist())

    # --------------------- true aoa measures --------------------- #
    true_clusters_evo <- clusters_evolution(vocab,
                                            n_clusters,
                                            min(vocab$age),
                                            max(vocab$age),
                                            randomization = "none")

    # w2v vocabulary sizes (number of words know at each month in clustering)
    measurements[["w2v_vocab_sizes"]] <- 
      tibble(age = min(vocab$age):max(vocab$age),
             measure = "w2v_vocab_size",
             randomization = NA, # same for all randomizations
             sample = NA,
             value = unlist(map(true_clusters_evo, ~ length(unlist(.)))))

    # dynamics measures
    for (measure in measures) {
      value <- run_measure(true_clusters_evo, measure)
      measurements[[measure]] <-
        tibble(age = min(vocab$age):max(vocab$age),
               measure = measure,
               randomization = "none",
               sample = NA,
               value = value)
    }

    # --------------------- randomized aoa measures --------------------- #
    for (randomization in randomizations) {

      for (sample in 1:samples) {
        random_clusters_evo <- clusters_evolution(vocab,
                                                  n_clusters,
                                                  min(vocab$age),
                                                  max(vocab$age),
                                                  randomization = randomization)

        for (measure in measures) {
          value <- run_measure(random_clusters_evo, measure)
          measurements[[paste(randomization, measure, sample, sep = "__")]] <-
            tibble(age = min(vocab$age):max(vocab$age),
                   measure = measure,
                   randomization = randomization,
                   sample = sample,
                   value = value)
        }
      }
    }

    return(bind_rows(measurements))
}

run_experiment <- function(languages, measures, randomizations, cluster_counts, samples) {
  measurements = list()
  cat("measuring dynamics\n")
  for(language in languages) {
    cat("  language:", language, "\n")
    for(n_clusters in cluster_counts) {
      cat("    n_clusters: ", n_clusters, "\n")
      measurements[[length(measurements) + 1]] <- 
        measure_cluster_dynamics(language, n_clusters, measures, randomizations, samples) %>%
        mutate(language = language) %>%
        mutate(n_clusters = n_clusters) %>%
        select(language, n_clusters, everything())
    }
  }
  return(bind_rows(measurements))
}
```

## Run Experiment

```{r warning=FALSE, message=FALSE}
languages <-  c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish")

measures <- c("pair_precision_to_last", "pair_recall_to_last", "pair_f1_to_last", "min_max_size_ratio")

randomizations <- c("random_aoa", "to_nearest_aoa", "within_last_clustering")

### compute measurements, TAKES A LONG TIME

# measurements <- run_experiment(languages = languages,
#                           measures = measures,
#                           randomizations = randomizations,
#                           cluster_counts = 2:5,
#                           samples = 50)
# 
#saveRDS(measurements, "measurements.rds")

measurements <- readRDS("measurements.rds")

```

## Summarize and Visualize Results
```{r}
to_pdf = FALSE

### Plot vocab sizes
vocab_sizes <- measurements %>%
  select(language, n_clusters, age, measure, value) %>%
  filter(measure %in% c("true_vocab_size", "w2v_vocab_size"))

if (to_pdf) {
  pdf("output/vocab_size.pdf", width=20, height=3)
}
ggplot(vocab_sizes) +
  geom_line(aes(x = age, y = value, colour = measure)) + 
  facet_grid(~ language)
if (to_pdf) {
  dev.off()
}
```
```{r}
### Plot measure summary
to_pdf = FALSE

for (measure_to_plot in measures) {
  for (randomization_to_plot in randomizations) {

    summary_fname <- paste(measure_to_plot, randomization_to_plot, sep="__") %>%
      paste("pdf", sep = ".")

    # APPROACH # 1
    true_evolution <- measurements %>%
      filter(measure == measure_to_plot) %>%
      filter(randomization == "none") %>%
      select(language, n_clusters, age, value)

    random_evolution <- measurements %>%
      filter(measure == measure_to_plot) %>%
      filter(randomization == randomization_to_plot) %>%
      group_by(language, n_clusters, age) %>%
      summarise(mean = mean(value),
                sd = sd(value),
                `25%`=quantile_func(value, probs=0.25),
                `50%`=quantile_func(value, probs=0.5),
                `75%`=quantile_func(value, probs=0.75),
                `10%`=quantile_func(value, probs=0.1),
                `90%`=quantile_func(value, probs=0.9)) %>%
      ungroup()

    evolution <- full_join(true_evolution, random_evolution, by = c("language", "n_clusters", "age"))

    if (to_pdf) {
      pdf(paste("output",summary_fname,sep="/"), width=20, height=10)
    }
    print(
      ggplot(evolution, aes(x = age)) +
        geom_ribbon(aes(ymin = `10%`,
                        ymax = `90%`),
                    colour="yellow",
                    fill="yellow",
                    linetype=2,
                    alpha= 0.2) +
        geom_ribbon(aes(ymin = `25%`,
                        ymax = `75%`),
                    colour="orange",
                    fill="orange",
                    linetype=2,
                    alpha= 0.3) +
        geom_line(aes(y = `50%`), color = "red", alpha = 0.4) + 
        geom_line(aes(y = value), color = "blue") +
        labs(title = paste("True value against",
                           randomization_to_plot,
                           "statistics on",
                           max(measurements$sample, na.rm = TRUE), 
                           "samples"),
             subtitle = "25%-75% and 10%-90% quantiles shown",
             x ="Age (m)", y = measure_to_plot) +
        facet_grid(n_clusters ~ language)
    )
    if (to_pdf) {
      dev.off()
    }
  }
}
```

# Deprecated

```{r}
# for pdf, put "pdf("precision_20_samples.pdf", width=10, height=5)" before and "dev.off()" after plotting
# ggplot(results_precision, aes(x=age)) +
#   geom_line(aes(y = full_vocab_size, colour = "Full Vocab")) +
#   geom_line(aes(y = w2v_vocab_size, colour = "W2V Vocab")) +
#   theme(legend.title=element_blank()) +
#   facet_grid(. ~ language) +
#   labs(title="Vocab size by month", x="Age", y="Vocab Size")

## plus-minus standard

for (measure in measures) {
  print(measure)
  # for (randomization in randomizations) {
  #   print("plotting")
  #   c("plotting\n")
  #   # mean_name <- random_measurement_name(measure, randomization, "mean")
  #   # sd_name <- random_measurement_name(measure, randomization, "sd")
  #   # 
  #   # print("plotting")
  #   # ggplot(results, aes(x=age)) +
  #   #   geom_line(aes(y = !! measure)) +
  #   #   facet_grid(n_clusters ~ language)
  # }
}

# ggplot(results_precision, aes(x=age)) +
#   geom_ribbon(aes(ymin=rand_measure_evo_mean - rand_measure_evo_sd,
#                   ymax=rand_measure_evo_mean + rand_measure_evo_sd),
#               colour="grey",
#               fill="grey",
#               linetype=2,
#               alpha= 0.2) +
#   geom_ribbon(aes(ymin=rand_measure_evo_mean - 2 * rand_measure_evo_sd,
#                   ymax=rand_measure_evo_mean + 2 * rand_measure_evo_sd),
#               colour="grey",
#               fill="grey",
#               linetype=2,
#               alpha= 0.2) +
# #  geom_line(aes(y = rand_measure_evo_mean),
# #            color="grey") +
#   geom_line(aes(y = measure_evo),
#             color="black") +
#   theme(legend.title=element_blank()) +
#   facet_grid(n_clusters ~ language) +
#   labs(title="Precision against clustering at 30 months", x="Age", y="Measure")

```