### 2019-02-12 conceptNet-cogsci/CogsciBackground.Rmd
Added CogsciBackground

---
title: "CogSci Background Analysis"
author: "Isaac Scheinfeld"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: default
---

# Setup

```{r echo=FALSE, message=FALSE, warning=FALSE} 
library(purrr) # fp library
library(readr) # read text data files (csv, etc.)
library(ggplot2) # plotting
library(dplyr) # for working with datafram-like objects
library(tidyr) # for data tidying
library(wordbankr) # access to wordbank data
library(stringr) # common string operations
library(igraph) # network analysis and visualization
library(tibble) # modern dataframes
library(tidygraph)
library(matrixStats)
library(memoise)

source("helpers/wb_helper.r")
source("helpers/word2vec_helper.r")
source("helpers/network_helper.r")
source("helpers/random_helper.r")
source("helpers/dynamics_helper.r")
source("helpers/statistics_helper.r")
```

## Nouns

We begin by importing data from wordbank. Age of aquisition (in months, it starts with 16, 17,..) is that at which 50% of children can produce the word. We also do naive triming at this point, this means we are ignoring homophone/polysemy (chiken (food), chiken (animals), etc,.. but we will have to be careful in our choices if we are preparing later for publication.

```{r warning=FALSE}
get_language_vocab_ <- function(lang) {
  return(make_vocab_dataframe(lang=lang,
                              lang_form = "WS",
                              lex_class = "nouns") %>%
           trim_all_unilemma() %>% # TODO check trimming
           trim_all_definition() %>%
           arrange(age) %>%            # keep first age for duplicates
           distinct(uni_lemma, .keep_all = TRUE) %>%
           select(uni_lemma, definition, age, category))
}

get_language_vocab <- memoize(get_language_vocab_) # memoize to prevent re-downloading dataset TODO check if this works
```

##

## Build Word2Vec Network Evolution
```{r message=FALSE}
test_vocab <- get_language_vocab("German")
test_w2v_network <- vocab_to_w2v_network(test_vocab)
test_feature_network <- vocab_to_feature_network(test_vocab, c('function'))

transitivity(test_w2v_network)
transitivity(test_feature_network)

# weights = E(test_w2v_network)$w2v_similarity
# weights = E(test_feature_network)$shared

```

```{r}
w2v_clusters <- cut_at(w2v_walktrap(test_w2v_network), no = 3)
feature_clusters <- cut_at(feature_walktrap(test_feature_network), no = 3)

plot_vocab_network(test_w2v_network, 
                   layout = NULL, 
                   name=NA, 
                   frame=FALSE, 
                   clusters=w2v_clusters, 
                   plot_title = NULL, 
                   edge_color = "grey", 
                   labels=V(test_w2v_network)$uni_lemma)
```