### 2019-02-12 CMCL/ms.Rmd
paper

---
title: "How to Make a Camera-Ready Proceedings Contribution"

author-information: > 
    \author{{\large \bf Author 1} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University
    \And {\large \bf Author 2} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Some University}

abstract: 
    "This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of EACL-2017. The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. These instructions should be
  used for both papers submitted for review and for final versions of
  accepted papers.  Authors are asked to conform to all the directions
  reported in this document."

final_submission: no
# If yes, uncomment line below and add your own ID
# eaclpaperid: "1234"

# Uncomment and change if you want a smaller titlebox; minimum length is 5cm
# titlebox_length: "5cm"

output: acl2017::acl_paper
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(dplyr)
library(ggplot2)
library(grid)
library(lme4)
library(lmerTest)
library(ggthemes)
library(ggplot2)
library(xtable)
```

# Methods

##Nouns 
We used data from Wordbank (Frank et al., 2017), an open repository aggregating cross-linguistic language developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist, Toddler version. Parent report is a reliable and valid measure of children's vocabulary (Fenson et al., 1994). We used a set of nouns representing hand-checked translation equivalents  across 10 languages (available in Wordbank database). We "assume" (i should ask what were the criteria for the unilemma selection) that this set of translation equivalents is representative of a core shared vocabulary across all children (in different cultures) by around 30 months, 
<!--and we study possible cross-lingusitic variation in the way this shared knowledge is arrived at in development. -->

The conceptual organization derived from this shared vocabulary is obviousely identical across languages, but we investigate whether there are systematic cross-lingusitc variations in the order of aquisition of words that make up this vocabulary, and crucially, whether such variation influences the induced conceptual organization at different points in time. In fact, differences in the order of acqusition of words do not necessarily give rise to different conceptual organization. Imagine that two languages vary in whether "cow" or "dog" is acquired first. This difference will not change the induced conceptual organization across time since both "dog" and "cow" are instances of the same high-level concept "animal".

```{r fig.env = "figure", fig.pos = "h", fig.align = "center", set.cap.width=T, fig.width=4, fig.height=3, num.cols.cap=2, fig.cap = "bla bla"}

cor_all <- feather::read_feather("../saved_data/data_corr_all.feather")

ggplot(cor_all, 
      aes(x = tau, fill=data)) +
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
                 alpha=0.5,
                 binwidth=0.05)+
  #scale_x_log10() +
  theme_few() + 
  theme(aspect.ratio = 0.7, legend.title = element_text(size=8)) +  #facet_grid(Segmentation ~ language)+
  scale_y_continuous(labels = scales::percent)+
  xlab("pairwise correlation") +ylab("Count")

```

## Features
We used two sources of information which represent two ways words may be related to one another in the semantic network. The first source was the MacRae features (McRae et al., 2005). These features were collected by giving adult participants a set of nouns and prompting them to provide various kinds of properties (perceptual, functional, encyclopedic, and taxonomic). We followed Hills et al. 2010 in excluding the encyclopedic and taxonomic properties as their role may not be important at this stage of development. The other source was the semantic similarity derived based on co-occurrence in the corpus of CHILDES, using word2vec.

##Network 
We constructed networks using CDI words as nodes. Pairs of words were linked by an edge weighted by the number of shared MacRae feature or with the continuous distance obtained via Word2vec.  To model change in the conceptual organization from month to month, we constructed a different network at each month, based on the words that have been acquired by that month. We defined the age of acquisiton of a given word by the month at which this noun was produced by at least 50% of children, in each language (Goodman et al., 2008).

## Small World properties
We test whether the networks display the so-called "small-world" properties similar to other semantic and real-world networks (Steyvers & Tenenbaum, 2005; Watts & Strogatz, 1998). Small world properties are characterized with the average clustering coefficient $C$ and the average shortest path $L$. The former measures the extent to which the network is clustered, i.e., made of highly connectned sub-networks, whereas the latter measures the typical separation between tow nodes in the network. A network is small-word if it has a higher clustering coeffient comapred to a randomly connected network of the same size $C \gg C_{random}$, while still having a shorest path length as small as the one typically observed in random networks, that is $L \approx L_{random}$.

```{r fig.env = "figure*", fig.pos = "h", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.width=7, fig.height=14, fig.cap = "bla bla"}
measurements <- readRDS("data/measurements.rds")

all <- measurements %>%
  filter(measure %in% c("pair_precision_to_last", "pair_recall_to_last"),
         randomization != "to_nearest_aoa") #%>%
  #group_by(n_clusters, age, measure, randomization) %>%
  #summarise(mean = mean(value))

all_sum <- all %>%
  group_by(measure, randomization, age) %>%
  summarise(myMean = mean(value, na.rm = TRUE))

ggplot(data=all, aes(x=age, y=value, col=randomization)) +
  #geom_point() +
  geom_smooth(method="lm", formula = y ~ poly(x, 3)) +
  #geom_smooth(method="lm", formula = y ~ log(x)) +
  theme_few() + 
  theme(aspect.ratio = 0.7)+
  facet_grid(.~measure)
```
## Clustering analysis
We identified network clusters using WalkTrap (Pons & Latapy, 2006), a community detection algorithm based on the fact that a random walker tends to be trapped in dense parts of a network corresponding to highly interconnected sub-groups of nodes.

## Evaluation of clusterings aross development
We note $\mathcal{C}_t$ the clustering at month $t$, based on the subset of the vocabulary acquired by that month in each language. The evaluation consists in comparing $\mathcal{C}_t$ to the final clustering $\mathcal{C}^*$ obtained using the full vocabulary by the last month of acquisition (i.e., $\mathcal{C}^*$=$\mathcal{C}_t$ for $t$=30 months). Such comparison allows us to study how words acquired at different points in time may induce changes in the conceptual organization. In addition, we investigate consistency and variability in this change across languages.

We compare $\mathcal{C}_t$ to $\mathcal{C}^*$ using a standard method in clustering comparison, which is based on counting word pairs on which the two clusterings agree or disagree (Rand, 1971; Hubert and Arabie, 1985). A pair of words learned by month $t$ can fall under one of the four follwoing cases: 1) pairs that are placed in the same cluster under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$ (True positives, noted as $tp(\mathcal{C}_t)$), 2) pairs placed in different clsuters under $\mathcal{C}_t$ and in different clusters under $\mathcal{C}^*$ (True negatives, $tn(\mathcal{C}_t)$), 3) pairs placed in the same cluster under $\mathcal{C}_t$ and in different clusters under $\mathcal{C}^*$ (False positive, $fp(\mathcal{C}_t)$), and 4) pairs placed in different clusters under $\mathcal{C}_t$ and in the same cluster under $\mathcal{C}^*$ (False negatives, $fn(\mathcal{C}_t)$).

The clustering precison $P(\mathcal{C}_t)$ and recall $R(\mathcal{C}_t)$ are defiend as follows:

$$
P(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fp(\mathcal{C}_t)|}
$$
$$
R(\mathcal{C}_t) = \frac{|tp(\mathcal{C}_t)|}{|tp(\mathcal{C}_t)| + |fn(\mathcal{C}_t)|}
$$
Both Precison and Recall converge to 1 (perfect score) as $\mathcal{C}_t$ becomes more and more similar to $\mathcal{C}^*$. If precision starts low before converging to 1 (as opposed to being a constant at 1), this pattern would indicate that some pairs that should be differentiated are initially lumped together, suggesting a process of "differentiation" over development. Similarly, if we observe an increase in Recall, this pattern would indicate that some pairs that should be associated are initially differentiate, suggesting a process of "coelescence" over development.

## Word ordering 
The question we are addressing is whether the order of acqusition of words affect the

#Results

For real and random: results show that both precision and recall increase across development, suggesting that the conceptual organization undergoes both differentiation for some word pairs (example), and coelescecne for other pairs (examples).

For within-cluster: precision is high, but recall is low: explain why

How to explain this change?

>Assuming a random/real ranking of words:
-Some words may be lamped together initially (even if we force a relatively high number of clusters) because of the smaller vocabulary size? Can we have an example of this? Are the other clsuters empty?
-Some words may be differentiated initially because of the forcing?
-Is there a roleof noise? maybe clustering is just more random for smaller size vocabulary?

>Assuming a with-cluster ranking of words:
-Some words may be lamped together initially (even if we force a relatively high number of clusters): this does not happen, at least while we are still working with the first cluster...that's why precision is generally high
-Some words may be differentiated initially because they

<!--
If precision is lower early on, it means that initial concepts are not differentiated enough, that is, pairs that should belong to different concpets are intially lamped together (false positives), despite the fact that we forced the same number of clsuters in both $\mathcal{C}_t$ and $\mathcal{C}^*$ (Why is this the case? This looks like an artefact of the clustering algorithm with smaller size data? Should discuss this with Isaac. Also, what is an exmaple of a clustering $\mathcal{C}_t$ for a smaller size: are some clusters totally empty?)

Quetions:
Why does the control precision decreese a bit during developmnt, isn't supposed to be 1 all along? since we are splitting words within the same cluster (unless maybe some clusters are empty) : Oh I guess because we start introducing additional categories
-->

## Order of word learning

# Sudy of the conceptual development

Models

```{r}
# What is the model that can predict values as a function of cluser number and the methio 
#m1_precision <- lmer(value ~ age *  randomization + (age *  randomization | language) + (1 | n_clusters),
#           data = subset(all, measure =='pair_precision_to_last' & randomization != "random_aoa"),
#          control=lmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

##Test of convergence
#(From: https://github.com/lme4/lme4/issues/120)
#relgrad <- with(m1_precision@optinfo$derivs,solve(Hessian,gradient))
#max(abs(relgrad))

#saveRDS(m1_precision, "model_1.rds")
m1_precision <- readRDS("model_1.rds")

#m2_precision <- lmer(value ~ age *  randomization + (age *  randomization | language) + (1 | n_clusters),
#           data = subset(all, measure =='pair_precision_to_last' & randomization != "within_last_clustering"),
#           control=lmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

#saveRDS(m2_precision, "model_2.rds")
m2_precision <- readRDS("model_2.rds")

##Test of convergence
#relgrad <- with(m2_precision@optinfo$derivs,solve(Hessian,gradient))
#max(abs(relgrad))

#m1_recall <- lmer(value ~ age *  randomization + (age *  randomization | language) + (1 | n_clusters),
#           data = subset(all, measure =='pair_recall_to_last' & randomization != "random_aoa"),
#          control=lmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

#saveRDS(m1_recall, "model_recall_1.rds")
m1_recall <- readRDS("model_recall_1.rds")

#m2_recall <- lmer(value ~ age *  randomization + (age *  randomization | language) + (1 | n_clusters),
#           data = subset(all, measure =='pair_recall_to_last' & randomization != "within_last_clustering"),
#          control=lmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

#saveRDS(m2_recall, "model_recall_2.rds")
m2_recall <- readRDS("model_recall_2.rds")

```

# Important citation stuff! READ!

## Why we can't have nice things

So because the ACL committee wants their `.tex` files all nice and consistent, the old version of this package isn't good for what they want.  The Markdown-to-Latex conversion would automatically add in the references and citations, but would literally hardcode them into the `.tex` file.  We can't have that, now can we?  The only ways you were allowed to cite in your `.tex` file was with `\cite{Gusfield97}`, `\shortcite{Gusfield97}`, or `\newcite{Gusfield97}`, which corresponded to "(Gusfield, 1997)", "(1997)", or "Gusfield (1997)", respectively.

This means you couldn't do anything like "(e.g., Gusfield, 2017)", and their `.bst` file was incompatible with `natbib`. Obviously.  However, I have come to the rescue.  In the `eacl2017.sty` file in this package, I have added my own bit of magic: the `\barecite{}` command, which corresponds to "Burchill, 2017".  Notice this version doesn't have parentheses, so you can get stuff like "(e.g., Burchill, 2017)" with `(e.g., \barecite{Gusfield97})`.  Just be careful about making sure you don't forget parentheses.  Also not that when you submit the `.zip` folder, you should make sure to include the edited `eacl2017.sty` file.

## Examples

 * `\cite{Gusfield97}` becomes \cite{Gusfield97}
 * `\shortcite{Gusfield97}` becomes \shortcite{Gusfield97}
 * `\newcite{Gusfield97}` becomes \newcite{Gusfield97}
 * `\barecite{Gusfield97}` becomes \barecite{Gusfield97}
 * `\cite{Gusfield97,Aho72}` becomes \cite{Gusfield97,Aho72}
 * `\barecite{Gusfield97,Aho72}` becomes \barecite{Gusfield97,Aho72}

**Wait, _are those citations appearing as question marks?_  Don't worry, read on!**

## The two-step process of knitting this file

Because the ACL's you-need-to-use-our-citing-function formatting doesn't work well with R Markdown--which really likes to compile its own citations--knitting this is now a two-step process.  After you get it how you want it in RMarkdown, knit as usual.  You'll see that all citations are question marks, and that the bibliography is missing.  This is natural, don't panic.

For the current purposes, let's assume this file we're in right now is called `acl_draft.Rmd`.  Navigate into the directory containing `acl_draft.Rmd` via a terminal.  Then run the following command (swapping out your actual file name for "acl_draft")

`pdflatex acl_draft.tex; pdflatex acl_draft.tex; bibtex acl_draft.aux; pdflatex acl_draft.tex; pdflatex acl_draft.tex`

Now, check the `.pdf` file.  If all went well, it should now have all the citations in it as well as the reference section.  Good luck!

# General Formatting Instructions

All the default content below is lifted directly from Kyle MacDonald's Cogsci 2016 project (https://github.com/kemacdonald/cogsci2016), which this entire package is based on.  I haven't changed it, so it's up to you to ignore as you please.

For general information about authoring in markdown, see **http://rmarkdown.rstudio.com/authoring_basics.html.**

For standard spoken papers and standard posters, the entire 
contribution (including figures, references, everything) can be 
no longer than six pages. For abstract posters, the entire contribution 
can be no longer than one page. For symposia, the entire contribution 
can be no longer than two pages.

The text of the paper should be formatted in two columns with an
overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5
cm), with 0.25 inches between the columns. Leave two line spaces
between the last author listed and the text of the paper. The left
margin should be 0.75 inches and the top margin should be 1 inch.
\textbf{The right and bottom margins will depend on whether you use
U.S. letter or A4 paper, so you must be sure to measure the width of
the printed text.} Use 10 point Times Roman with 12 point vertical
spacing, unless otherwise specified.

The title should be in 14 point, bold, and centered. The title should
be formatted with initial caps (the first letter of content words
capitalized and the rest lower case). Each author's name should appear
on a separate line, 11 point bold, and centered, with the author's
email address in parentheses. Under each author's name list the
author's affiliation and postal address in ordinary 10 point  type.

Indent the first line of each paragraph by 1/8~inch (except for the
first paragraph of a new section). Do not add extra vertical space
between paragraphs.

# First-Level Headings

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and 1/4~line space
below the heading.

## Second-Level Headings

Second level headings should be 11 point , initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

### Third-Level Headings

Third-level headings should be 10 point , initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

# Formalities, Footnotes, and Floats

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.}

## Figures

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you can change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", fig.cap = "This image spans both columns."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```

## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
       aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                      caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

```{r}
# References will be generated after this point after you run the following command. <file> is this file, but without the extension.  You should include the full path if you aren't running the command in the same directory.  This command should be run in a terminal, btw.

# The command:
# pdflatex <file>.tex; pdflatex <file>.tex; bibtex <file>.aux; pdflatex <file>.tex; pdflatex <file>.tex

# After running this command, check the new pdf file. It should hopefully have all the citations added in.
```